\bf Comparing the Power of Plot Designs to Reveal Correlation

### 2016-01-11 paper/paper.Rnw
changes
TODO and fig refs

\begin{figure}[htbp]
  \centering
<<designs, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
set.seed(101)
a <- gen_true_data(24, 0.95)
b <- gen_true_data(24, -0.95)
a_l <- gather(a, variable, value, -t)
b_l <- gather(b, variable, value, -t)
p1b <- ggplot(data=a_l, aes(x=t, y=value, colour=variable)) + geom_line() + labs(title="Overlaid line graph") +
      theme(axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          legend.position="none")
p2b <- ggplot(data=b_l, aes(x=t, y=value, colour=variable)) + geom_line() + labs(title="Overlaid line graph") +
      theme(axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          legend.position="none")
p3b <- ggplot(data=a, aes(x=X1, y=X2)) + geom_point() + labs(title="Scatterplot") +
      theme(aspect.ratio=1, axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          legend.position="none")
p4b <- ggplot(data=b, aes(x=X1, y=X2)) + geom_point() + labs(title="Scatterplot") +
      theme(aspect.ratio=1, axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          legend.position="none")
grid.arrange(p1b, p2b, p3b, p4b, ncol=2)
@
  \label{designs}
  \caption{The same two data sets in both plot designs. t=24, correlation (left) = 0.946, correlation (right) = -0.943}
\end{figure}

@@

\begin{figure}[htbp]
 \centering
<<results, dependson='setup', fig.width=6, fig.height=8, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, error=FALSE, message=FALSE>>=
Turk <- mutate(Turk, plot = ifelse(test_param =="line", "Overlaid Lines", "Scatterplot"), 
               sign=ifelse(correlation > 0, 1, 0), smoothed=ifelse(smoothed=="yes", 1, 0))
#ggplot(data=subset(Turk, smoothed==0), aes(x=correlation, y=percent, colour=plot)) + geom_smooth() + geom_point() + facet_grid(n~.) + scale_colour_discrete(name = "Plot Design") +  labs(title="Unsmoothed Plot Accuracy", y="Percent Correct", x="True Plot Correlation")
ggplot(data=Turk, aes(x=correlation, y=percent, colour=plot)) + geom_smooth(se=F) + geom_point() + facet_grid(n~smoothed) + 
  scale_colour_discrete(name = "Plot Design") + labs(y="True plot detection rate", x="Correlation") +
  theme(legend.position="bottom")
@
  \label{results}
  \caption{Response accuracy for unsmoothed data, broken by unsmoothed (left), smoothed (right) and sample size. The dots are data points, the lines represent a naive model average fit. Both plot designs have similar performance for positively correlated data, however the overlaid line graphs are particularly weak at revealing negative correlation.}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
<<powerdiff, dependson='results', fig.width=6, fig.height=8, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
Turk$correct <- Turk$percent*Turk$attempt
Turk$r <- round(Turk$correlation, 1)
#Turk.line <- subset(Turk, test_param=="line")
#Turk.scatter <- subset(Turk, test_param=="scatter")
prop <- Turk[order(Turk$r, Turk$n, Turk$smoothed),c("smoothed", "n", "attempt", "correct", "r", "plot")]
prop <- summarise(group_by(prop, smoothed, n, r), l.attempt=sum(ifelse(plot=="Overlaid Lines", attempt, 0)),
                                                  l.correct=sum(ifelse(plot=="Overlaid Lines", correct, 0)),
                                                  s.attempt=sum(ifelse(plot=="Scatterplot", attempt, 0)),
                                                  s.correct=sum(ifelse(plot=="Scatterplot", correct, 0)))
#Turk.line <- Turk.line[order(Turk.line$r, Turk.line$n, Turk.line$smoothed),c("smoothed", "n", "attempt", "correct", "r")]
#Turk.line <- summarise(group_by(Turk.line, smoothed, n, r, test_param), sum(attempt), sum(correct))
#Turk.scatter <- Turk.scatter[order(Turk.scatter$r, Turk.scatter$n, Turk.scatter$smoothed),c("smoothed", "n", "attempt", "correct", "r")]
#Turk.scatter <- summarise(group_by(Turk.scatter, smoothed, n, r), sum(attempt), sum(correct))
#colnames(Turk.scatter) <- c("smoothed", "n", "r", "s.attempt", "s.correct")
#colnames(Turk.line) <- c("smoothed", "n", "r", "l.attempt", "l.correct")
#prop <- cbind(Turk.line, Turk.scatter[4:5])
prop <- mutate(prop, l.adj = (l.correct+1)/(l.attempt+1), s.adj = (s.correct+1)/(s.attempt+1),
               df = s.attempt + l.attempt - 1, difference=s.adj-l.adj, t = qt(0.975, df),
               upper = difference + t*sqrt(s.adj*(1-s.adj)/s.attempt+l.adj*(1-l.adj)/l.attempt),
               lower = difference - t*sqrt(s.adj*(1-s.adj)/s.attempt+l.adj*(1-l.adj)/l.attempt))
ggplot(data=prop) + geom_errorbar(aes(x=r, ymax=upper, ymin=lower)) + geom_point(aes(x=r, y=difference)) +
        geom_hline(aes(yint=0, colour="red")) + labs(x="Correlation", y="Scatterplot power minus Overlaid Lines") + facet_grid(n~smoothed)
@
 \label{powerdiff}
 \caption{The difference in estimated power (scatter plot power minus overlaid line power) to reveal the true plot by smoothness and sample size with a 95\% confidence interval. The red line indicates the point where both plot designs are equally powerful while values above this line indicate that the scatter plot is more powerful than the overlaid line graph. The largest differences appear in negative correlation, where the scatter plot is more powerful.}
\end{figure}

\begin{figure}[htbp]
  \centering
<<samplesize, dependson='powerdiff', fig.width=6, fig.height=8, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
ggplot(data=subset(Turk, smoothed==0), aes(x=correlation, y=percent, colour=factor(n))) + geom_smooth(se=F) + geom_point() + facet_grid(plot~.) + scale_colour_discrete(name = "Sample size") +  labs(y="Percent Correct", x="Correlation") + theme(legend.position="bottom")
@
 \label{samplesize}
 \caption{Replotting figure~\ref{results} to highlight the effects of changes in sample size of unsmoothed data. The scatter plot appears to universaly benefit from increased sample sizes but the overlaid lines only see a benefit in positive correlation.}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
<<time, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
full <- mutate(full, sign = ifelse(correlation >= 0, 1, 0), n24 = ifelse(n ==24, 1, 0), n48 = ifelse(n ==48,  1, 0), n96 = ifelse(n ==96, 1, 0), plot = ifelse(test_param ==1, "Overlaid Lines", "Scatterplot"))
time <- data.frame(plot=full$plot, time=full$log_time, correct=full$correct, rsgn=full$sign)
#ggplot(data=time, aes(x=time, fill=factor(correct))) +geom_bar(stat="bin") + facet_grid(~plot) +labs(x="Log(Time) Taken", y=element_blank()) + scale_fill_discrete(name = "Correct")
time$rsgn <- factor(time$rsgn, labels=c("r -","r +"))
tm <- summarise(group_by(time, correct, plot, rsgn), m=mean(time))
t1 <- ggplot(data=subset(time, correct==0), aes(x=time)) + geom_histogram(binwidth=0.5) + 
  geom_vline(data=subset(tm, correct==0), aes(xintercept = m), color="red") +
  facet_grid(plot~rsgn) + labs(title="Did not detect", x="Log(Time) Taken", y=element_blank())  
t2 <- ggplot(data=subset(time, correct==1), aes(x=time)) + geom_histogram(binwidth=0.5) + 
  geom_vline(data=subset(tm, correct==1), aes(xintercept = m), color="red") +
  facet_grid(plot~rsgn) + labs(title="Detected", x="Log(Time) Taken", y=element_blank())  
grid.arrange(t1, t2, ncol=2)
@
  \label{time} 
  \caption{Histograms of log time taken to make a decision. r+ and r- indicate the sign of true plot correlation. The vertical red line is the average time of the particular group. On average scatter plots are faster and the true plot is detected more often, with r-.}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
<<confidence, dependson='time', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
conf <- data.frame(plot=full$plot, conf=full$conf_level, correct=full$correct, rsgn=full$sign)
conf$rsgn <- factor(conf$rsgn, labels=c("r -","r +"))
cm <- summarise(group_by(conf, correct, plot, rsgn), m=mean(conf))
c1 <- ggplot(data=subset(conf, correct==0), aes(x=conf)) + geom_histogram(binwidth=1) + 
  geom_vline(data=subset(cm, correct==0), aes(xintercept = m), color="red") +
  facet_grid(plot~rsgn) + labs(title="Did not detect", x="Confidence", y=element_blank())  
c2 <- ggplot(data=subset(conf, correct==1), aes(x=conf)) + geom_histogram(binwidth=1) + 
  geom_vline(data=subset(cm, correct==1), aes(xintercept = m), color="red") +
  facet_grid(plot~rsgn) + labs(title="Detected", x="Confidence", y=element_blank())  
grid.arrange(c1, c2, ncol=2)
@
  \label{confidence} 
  \caption{Histograms of reported confidence levels. 1 - Low, 5 - High. Detection had a minor detrimental impact on confidence, whilst time taken increased it. There was no statistically significant difference on confidence level between plot designs.}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
<<confidence2, dependson='time', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
full$subj_id <- factor(full$ip_address, labels=1:207)
y <- data.frame(table(full$subj_id))
keep2 <- y$Var1[y$Freq>9]
conf2 <- filter(full, subj_id %in% keep2)
conf2 <- select(conf2, conf_level, correlation, test_param, smoothed, n, correct, sign, plot, subj_id) #c(7, 15, 16, 17, 20, 27)]
#sub <- summarise(group_by(sub, subj_id), correct = sum(correct), conf=mean(conf_level))
#sub <- merge(sub, y, by.x="subj_id", by.y="Var1")
#sub <- mutate(sub, prop = correct/Freq)
qplot(correlation, 6-conf_level, data=conf2, alpha=I(0.1)) + 
  geom_smooth(method="loess") + 
  facet_grid(n~plot+smoothed) +
  labs(x = "Correlation", y = "Confidence (5=High, 1=Low)")
@
  \label{confidence2}
  \caption{Reported confidence against true plot correlation. Darker points indicate a confidence level selected more regularly. Strong correlation is more often associated with a lower reported confidence.}
\end{figure}

@@

  \label{pins}
  \caption{Pin plots of the subject selections of each lineup, by design, smoothness, true plot correlaition and sample size. On the x-axis is the correlation of the plot (there is potentially spurious correlation in the null plots), and on the y-axis is the proportion a plot was chosen. Each plot contains data from the three lineups with the requisite characteristics. Red pins indicate null plots and blue indicates the true plot, occasionally the true plot does not seem to appear as it has a proportion of zero and is visually blocked by the red pins. Strongly correlated plot seem to be selected regardless of if the relationship is real or spurious.}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.5\textwidth]{anom1}
  \label{anom1}
  \caption{A lineup of unsmoothed scatter plots with a sample size of 24. The true plot is located in position 4 with a correlation of -0.491, and was selected by one out of eleven observers (p = 0.33). The plot in position 6 was selected by nine of these eleven observers (p < 0.001) despite having a correlation of -0.037. Subjects most often cited the clustering effect shown as the reason for their selection, an example of the over-interpretation of randomness.}
\end{figure}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.5\textwidth]{anom2}
  \label{anom2}
  \caption{A lineup of unsmoothed overlaid lines with a sample size of 12 and true correlation of -0.707. The true plot in position 8 was not detected in five evaluations. However, the null plot in position 13 was selected four times, (p –value < 0.001) despite having a correlation of only 0.225. When asked for a reason of their selections, subjects typically answered that the lines were close together, another misleading visual artefact of the random data.} 
\end{figure}

@@

\begin{eqnarray} 
\label{modelpower}
Power_{ijklt} &=& \psi \{ \beta_{0} + \beta_{1} x_{1,i}  + \beta_{2} x_{2,j}  + \beta_{3} x_{3,k} + \beta_{4} x_{4,t} + \\
 & & \beta_{5} x_{1,i} \* x_{2,j} + \beta_{6} x_{5,l}(\beta_{7} + \beta_{8} x_{1,i}  + \beta_{9} x_{2,j}  + \beta_{10} x_{3,k} + \nonumber \\
 & & \beta_{11} x_{4,t} + \beta_{12} x_{1,i} \* x_{2,j}) \gamma z_{s} + \epsilon_{ijklt} \} \nonumber
\end{eqnarray}

@@

qplot(r, RandomEf, data=prediction, color=plot, 
      group=interaction(subj_id, plot), geom="line", alpha=I(0.1)) + 
  scale_color_manual("Plot Design", values=c("#33AA33", "#9977EE", "#FF9933", "#00DDAA")) + 
  xlab("Correlation") + ylab("Estimated Power") +
  geom_line(aes(y=FixedEf), size=2, alpha=1) + theme_bw() + facet_wrap(~t, ncol = 2) + theme(aspect.ratio=1.2)
@
  \label{powermodel}
  \caption{Predicted power fit from the mixed effects logit power model. The thicker lines shows the fixed effects prediction while the thinner lines show the range of random effects, the observer's visual ability. The weakness of the overlaid line graph in negative correlation and the scatter plot's gain from sample size are highlighted.}
\end{figure}

\begin{figure}
  \centering
<<residuals, dependson='powermodel', fig.height=4, fig.width=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
ggplot(data=glmm.resid) + geom_histogram(aes(y=..density.., x=residual), binwidth=2) + labs(x=NULL, y=NULL)
@
  \label{residuals}
  \caption{Histograms of residuals, $\hat{\epsilon}$ for both the scatter plot (left) and the overlaid lines (right). Both appear to be normal, though there are some extreme outliers.}
\end{figure}

@@

\begin{figure}[htbp]
\label{random}
  \centering
<<random, dependson='powerfit', fig.height=4, fig.width=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
ggplot(data=randomeffects) + geom_histogram(aes(x=re, y=..density..), binwidth=0.08) + xlab("Random Effects") + ylab(NULL)
@
    \caption{Histograms of estimated subject random effects, $\hat{\gamma}$}
\end{figure}

@@

\begin{figure}[htbp]
\centering
\label{random2}
<<random2, dependson='random', fig.height=6, fig.width=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
turk.correct <- summarise(group_by(full.sub, pic_id), correct=sum(correct), attempts=sum(weights*attempts))
turk.correct$percent <- turk.correct$correct/turk.correct$attempts
ability <- merge(full.sub[,c("subj_id", "pic_id", "correct")], turk.correct[c("pic_id", "percent")], by="pic_id")
ability <- summarise(group_by(ability, subj_id), actual=sum(correct), estimated=sum(percent))
ability$diff <- ability$actual - ability$estimated
ability.re <- merge(ability, randomeffects, by="subj_id")
ggplot(data=ability.re, aes(x=diff, y=re)) + geom_point() + labs(x="Visual Ability", y="Random Effects")
@
\caption{Visual ability plotted against random effects. The diagonal pattern suggests a strong positive relationship, indicating that the random effects in the model are a suitable measure of visual ability.}
\end{figure}

@@

\begin{table}
\label{tab2}
\centering
\caption{Linear mixed model fit for time taken to make a selection, default plot style is overlaid lines. Fitted with the MIXED procedure in SAS.}

>>>>

\begin{figure}[htbp]
  \centering
<<designs, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
set.seed(101)
a <- gen_true_data(24, 0.95)
b <- gen_true_data(24, -0.95)
a_l <- gather(a, variable, value, -t)
b_l <- gather(b, variable, value, -t)
p1b <- ggplot(data=a_l, aes(x=t, y=value, colour=variable)) + geom_line() + labs(title="Overlaid line graph") +
      theme(axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          legend.position="none")
p2b <- ggplot(data=b_l, aes(x=t, y=value, colour=variable)) + geom_line() + labs(title="Overlaid line graph") +
      theme(axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          legend.position="none")
p3b <- ggplot(data=a, aes(x=X1, y=X2)) + geom_point() + labs(title="Scatterplot") +
      theme(aspect.ratio=1, axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          legend.position="none")
p4b <- ggplot(data=b, aes(x=X1, y=X2)) + geom_point() + labs(title="Scatterplot") +
      theme(aspect.ratio=1, axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          legend.position="none")
grid.arrange(p1b, p2b, p3b, p4b, ncol=2)
@
  \caption{The same two data sets in both plot designs. t=24, correlation (left) = 0.946, correlation (right) = -0.943}
    \label{designs}
\end{figure}

@@

\begin{figure}[htbp]
 \centering
<<results, dependson='setup', fig.width=6, fig.height=8, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, error=FALSE, message=FALSE>>=
Turk <- mutate(Turk, plot = ifelse(test_param =="line", "Overlaid Lines", "Scatterplot"), 
               sign=ifelse(correlation > 0, 1, 0), smoothed=ifelse(smoothed=="yes", 1, 0))
#ggplot(data=subset(Turk, smoothed==0), aes(x=correlation, y=percent, colour=plot)) + geom_smooth() + geom_point() + facet_grid(n~.) + scale_colour_discrete(name = "Plot Design") +  labs(title="Unsmoothed Plot Accuracy", y="Percent Correct", x="True Plot Correlation")
ggplot(data=Turk, aes(x=correlation, y=percent, colour=plot)) + geom_smooth(se=F) + geom_point() + facet_grid(n~smoothed) + 
  scale_colour_discrete(name = "Plot Design") + labs(y="True plot detection rate", x="Correlation") +
  theme(legend.position="bottom")
@
  \caption{Response accuracy for unsmoothed data, broken by unsmoothed (left), smoothed (right) and sample size. The dots are data points, the lines represent a naive model average fit. Both plot designs have similar performance for positively correlated data, however the overlaid line graphs are particularly weak at revealing negative correlation.}
    \label{results}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
<<powerdiff, dependson='results', fig.width=6, fig.height=8, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
Turk$correct <- Turk$percent*Turk$attempt
Turk$r <- round(Turk$correlation, 1)
#Turk.line <- subset(Turk, test_param=="line")
#Turk.scatter <- subset(Turk, test_param=="scatter")
prop <- Turk[order(Turk$r, Turk$n, Turk$smoothed),c("smoothed", "n", "attempt", "correct", "r", "plot")]
prop <- summarise(group_by(prop, smoothed, n, r), l.attempt=sum(ifelse(plot=="Overlaid Lines", attempt, 0)),
                                                  l.correct=sum(ifelse(plot=="Overlaid Lines", correct, 0)),
                                                  s.attempt=sum(ifelse(plot=="Scatterplot", attempt, 0)),
                                                  s.correct=sum(ifelse(plot=="Scatterplot", correct, 0)))
#Turk.line <- Turk.line[order(Turk.line$r, Turk.line$n, Turk.line$smoothed),c("smoothed", "n", "attempt", "correct", "r")]
#Turk.line <- summarise(group_by(Turk.line, smoothed, n, r, test_param), sum(attempt), sum(correct))
#Turk.scatter <- Turk.scatter[order(Turk.scatter$r, Turk.scatter$n, Turk.scatter$smoothed),c("smoothed", "n", "attempt", "correct", "r")]
#Turk.scatter <- summarise(group_by(Turk.scatter, smoothed, n, r), sum(attempt), sum(correct))
#colnames(Turk.scatter) <- c("smoothed", "n", "r", "s.attempt", "s.correct")
#colnames(Turk.line) <- c("smoothed", "n", "r", "l.attempt", "l.correct")
#prop <- cbind(Turk.line, Turk.scatter[4:5])
prop <- mutate(prop, l.adj = (l.correct+1)/(l.attempt+1), s.adj = (s.correct+1)/(s.attempt+1),
               df = s.attempt + l.attempt - 1, difference=s.adj-l.adj, t = qt(0.975, df),
               upper = difference + t*sqrt(s.adj*(1-s.adj)/s.attempt+l.adj*(1-l.adj)/l.attempt),
               lower = difference - t*sqrt(s.adj*(1-s.adj)/s.attempt+l.adj*(1-l.adj)/l.attempt))
ggplot(data=prop) + geom_errorbar(aes(x=r, ymax=upper, ymin=lower)) + geom_point(aes(x=r, y=difference)) +
        geom_hline(aes(yint=0, colour="red")) + labs(x="Correlation", y="Scatterplot power minus Overlaid Lines") + facet_grid(n~smoothed)
@
 \caption{The difference in estimated power (scatter plot power minus overlaid line power) to reveal the true plot by smoothness and sample size with a 95\% confidence interval. The red line indicates the point where both plot designs are equally powerful while values above this line indicate that the scatter plot is more powerful than the overlaid line graph. The largest differences appear in negative correlation, where the scatter plot is more powerful.}
 \label{powerdiff}
\end{figure}

\begin{figure}[htbp]
  \centering
<<samplesize, dependson='powerdiff', fig.width=6, fig.height=8, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
ggplot(data=subset(Turk, smoothed==0), aes(x=correlation, y=percent, colour=factor(n))) + geom_smooth(se=F) + geom_point() + facet_grid(plot~.) + scale_colour_discrete(name = "Sample size") +  labs(y="Percent Correct", x="Correlation") + theme(legend.position="bottom")
@
 \caption{Replotting figure~\ref{results} to highlight the effects of changes in sample size of unsmoothed data. The scatter plot appears to universaly benefit from increased sample sizes but the overlaid lines only see a benefit in positive correlation.}
 \label{samplesize}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
<<time, dependson='setup', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
full <- mutate(full, sign = ifelse(correlation >= 0, 1, 0), n24 = ifelse(n ==24, 1, 0), n48 = ifelse(n ==48,  1, 0), n96 = ifelse(n ==96, 1, 0), plot = ifelse(test_param ==1, "Overlaid Lines", "Scatterplot"))
time <- data.frame(plot=full$plot, time=full$log_time, correct=full$correct, rsgn=full$sign)
#ggplot(data=time, aes(x=time, fill=factor(correct))) +geom_bar(stat="bin") + facet_grid(~plot) +labs(x="Log(Time) Taken", y=element_blank()) + scale_fill_discrete(name = "Correct")
time$rsgn <- factor(time$rsgn, labels=c("r -","r +"))
tm <- summarise(group_by(time, correct, plot, rsgn), m=mean(time))
t1 <- ggplot(data=subset(time, correct==0), aes(x=time)) + geom_histogram(binwidth=0.5) + 
  geom_vline(data=subset(tm, correct==0), aes(xintercept = m), color="red") +
  facet_grid(plot~rsgn) + labs(title="Did not detect", x="Log(Time) Taken", y=element_blank())  
t2 <- ggplot(data=subset(time, correct==1), aes(x=time)) + geom_histogram(binwidth=0.5) + 
  geom_vline(data=subset(tm, correct==1), aes(xintercept = m), color="red") +
  facet_grid(plot~rsgn) + labs(title="Detected", x="Log(Time) Taken", y=element_blank())  
grid.arrange(t1, t2, ncol=2)
@
  \caption{Histograms of log time taken to make a decision. r+ and r- indicate the sign of true plot correlation. The vertical red line is the average time of the particular group. On average scatter plots are faster and the true plot is detected more often, with r-.}
  \label{time} 
\end{figure}

@@

\begin{figure}[htbp]
  \centering
<<confidence, dependson='time', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
conf <- data.frame(plot=full$plot, conf=full$conf_level, correct=full$correct, rsgn=full$sign)
conf$rsgn <- factor(conf$rsgn, labels=c("r -","r +"))
cm <- summarise(group_by(conf, correct, plot, rsgn), m=mean(conf))
c1 <- ggplot(data=subset(conf, correct==0), aes(x=conf)) + geom_histogram(binwidth=1) + 
  geom_vline(data=subset(cm, correct==0), aes(xintercept = m), color="red") +
  facet_grid(plot~rsgn) + labs(title="Did not detect", x="Confidence", y=element_blank())  
c2 <- ggplot(data=subset(conf, correct==1), aes(x=conf)) + geom_histogram(binwidth=1) + 
  geom_vline(data=subset(cm, correct==1), aes(xintercept = m), color="red") +
  facet_grid(plot~rsgn) + labs(title="Detected", x="Confidence", y=element_blank())  
grid.arrange(c1, c2, ncol=2)
@
  \caption{Histograms of reported confidence levels. 1 - Low, 5 - High. Detection had a minor detrimental impact on confidence, whilst time taken increased it. There was no statistically significant difference on confidence level between plot designs.}
  \label{confidence} 
\end{figure}

@@

\begin{figure}[htbp]
  \centering
<<confidence2, dependson='time', fig.width=6, fig.height=6, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
full$subj_id <- factor(full$ip_address, labels=1:207)
y <- data.frame(table(full$subj_id))
keep2 <- y$Var1[y$Freq>9]
conf2 <- filter(full, subj_id %in% keep2)
conf2 <- select(conf2, conf_level, correlation, test_param, smoothed, n, correct, sign, plot, subj_id) #c(7, 15, 16, 17, 20, 27)]
#sub <- summarise(group_by(sub, subj_id), correct = sum(correct), conf=mean(conf_level))
#sub <- merge(sub, y, by.x="subj_id", by.y="Var1")
#sub <- mutate(sub, prop = correct/Freq)
qplot(correlation, 6-conf_level, data=conf2, alpha=I(0.1)) + 
  geom_smooth(method="loess") + 
  facet_grid(n~plot+smoothed) +
  labs(x = "Correlation", y = "Confidence (5=High, 1=Low)")
@
  \caption{Reported confidence against true plot correlation. Darker points indicate a confidence level selected more regularly. Strong correlation is more often associated with a lower reported confidence.}
  \label{confidence2}
\end{figure}

@@

  \caption{Pin plots of the subject selections of each lineup, by design, smoothness, true plot correlaition and sample size. On the x-axis is the correlation of the plot (there is potentially spurious correlation in the null plots), and on the y-axis is the proportion a plot was chosen. Each plot contains data from the three lineups with the requisite characteristics. Red pins indicate null plots and blue indicates the true plot, occasionally the true plot does not seem to appear as it has a proportion of zero and is visually blocked by the red pins. Strongly correlated plot seem to be selected regardless of if the relationship is real or spurious.}
  \label{pins}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.5\textwidth]{anom1}
  \caption{A lineup of unsmoothed scatter plots with a sample size of 24. The true plot is located in position 4 with a correlation of -0.491, and was selected by one out of eleven observers (p = 0.33). The plot in position 6 was selected by nine of these eleven observers (p < 0.001) despite having a correlation of -0.037. Subjects most often cited the clustering effect shown as the reason for their selection, an example of the over-interpretation of randomness.}
  \label{anom1}
\end{figure}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.5\textwidth]{anom2}
  \caption{A lineup of unsmoothed overlaid lines with a sample size of 12 and true correlation of -0.707. The true plot in position 8 was not detected in five evaluations. However, the null plot in position 13 was selected four times, (p –value < 0.001) despite having a correlation of only 0.225. When asked for a reason of their selections, subjects typically answered that the lines were close together, another misleading visual artefact of the random data.} 
  \label{anom2}
\end{figure}

@@

\begin{eqnarray} 
\label{modelpower}
Power_{ijklt} &=& \psi \{ \beta_{0} + \beta_{1} x_{1,i}  + \beta_{2} x_{2,j}  + \beta_{3} x_{3,k} + \beta_{4} x_{4,t} + \\
 & & \beta_{5} x_{1,i} \* x_{2,j} + \beta_{6} x_{5,l}(\beta_{7} + \beta_{8} x_{1,i}  + \beta_{9} x_{2,j}  + \beta_{10} x_{3,k} + \nonumber \\
 & & \beta_{11} x_{4,t} + \beta_{12} x_{1,i} \* x_{2,j}) + \gamma z_{s} + \epsilon_{ijklt} \} \nonumber
\end{eqnarray}

@@

qplot(r, RandomEf, data=prediction, color=plot, 
      group=interaction(subj_id, plot), geom="line", alpha=I(0.1)) + 
  scale_color_manual("Plot Design", values=c("#33AA33", "#9977EE", "#FF9933", "#00DDAA")) + 
  xlab("Correlation") + ylab("Estimated Power") +
  geom_line(aes(y=FixedEf), size=2, alpha=1) + theme_bw() + facet_wrap(~t, ncol = 2) + theme(aspect.ratio=1.2)
@
  \caption{Predicted power fit from the mixed effects logit power model. The thicker lines shows the fixed effects prediction while the thinner lines show the range of random effects, the observer's visual ability. The weakness of the overlaid line graph in negative correlation and the scatter plot's gain from sample size are highlighted.}
  \label{powermodel}
\end{figure}

\begin{figure}
  \centering
<<residuals, dependson='powermodel', fig.height=4, fig.width=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
ggplot(data=glmm.resid) + geom_histogram(aes(y=..density.., x=residual), binwidth=2) + labs(x=NULL, y=NULL)
@
  \caption{Histograms of residuals, $\hat{\epsilon}$ for both the scatter plot (left) and the overlaid lines (right). Both appear to be normal, though there are some extreme outliers.}
  \label{residuals}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
<<random, dependson='powerfit', fig.height=4, fig.width=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
ggplot(data=randomeffects) + geom_histogram(aes(x=re, y=..density..), binwidth=0.08) + xlab("Random Effects") + ylab(NULL)
@
    \caption{Histograms of estimated subject random effects, $\hat{\gamma}$. They appear to be normally distributed with mean zero.}
\label{random}
\end{figure}

@@

\begin{figure}[htbp]
\centering
<<random2, dependson='random', fig.height=6, fig.width=6, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, message=FALSE, error=FALSE>>=
turk.correct <- summarise(group_by(full.sub, pic_id), correct=sum(correct), attempts=sum(weights*attempts))
turk.correct$percent <- turk.correct$correct/turk.correct$attempts
ability <- merge(full.sub[,c("subj_id", "pic_id", "correct")], turk.correct[c("pic_id", "percent")], by="pic_id")
ability <- summarise(group_by(ability, subj_id), actual=sum(correct), estimated=sum(percent))
ability$diff <- ability$actual - ability$estimated
ability.re <- merge(ability, randomeffects, by="subj_id")
ggplot(data=ability.re, aes(x=diff, y=re)) + geom_point() + labs(x="Visual Ability", y="Random Effects")
@
\caption{Visual ability plotted against random effects. The diagonal pattern suggests a strong positive relationship, indicating that the random effects in the model are a suitable measure of visual ability.}
\label{random2}
\end{figure}

\begin{table}[htbp]
\caption{Linear mixed model fit for time taken to make a selection, default plot style is overlaid lines. Fitted with the MIXED procedure in SAS.}
\centering
\label{tab2}
<<timefit, dependson='powerfit', echo=FALSE, results='hide', message=FALSE, error=FALSE, warning=FALSE>>=
#timelmer <-  lmer(log_time ~ smoothed + r + n + rsgn*r + plot + plot*smoothed 
#                  + plot*r + plot*n + plot*rsgn*r + plot*correct + (1|subj_id), data=full.sub)
#time.fit <- summary(timelmer)$coefficients
#pv <- function(x){
#  p <- 2*(1-pt(x, 1835))
#}
#p.value <- sapply(time.fit[,3], pv)
#time.fit <- cbind(time.fit, p.value)
time.fit <- read.csv("sas/time.csv")
time.fit <- time.fit[c(1, 5, 6, 2, 3, 4, 7, 8, 12, 13, 9, 10, 11, 14),] #separate into main effects then interaction effects in same order
rownames(time.fit) <- c("Intercept", "Smoothed", "t", "abs(r)", "sign(r)", "abs(r)*sign(r)", "detected", "Plot", "Plot*Smoothed", "Plot*t", "Plot*abs(r)", "Plot*sign(r)", "Plot*abs(r)*sign(r)", "Plot*detected")
time.fit$Effect <- NULL
colnames(time.fit) <- c("Estimate", "standard error", "df", "t-value", "p-value")
kable(time.fit[1:14,], format="latex", digits=2)
#kable(time.fit[8:14,], format="latex", caption="Interaction Effects", digits=2)
@
\end{table}

@@
